---
title: "assignment_5"
author: "Iris"
date: "November 7, 2014"
output: html_document
---
death probabilities

```{r, eval=TRUE}
delta.c <- 1/3
delta.r <- 10/32
tau <- 3/4
delta.s0 <- 1/4
delta.s <- delta.s0 + tau*fc
```

 put out initial outlay - start at equal abundances (from Fig.1 line graphs)
 
 0 = empty
 1=C
 2=R
 3=S
 
```{r}
possible.states <- c(0, 1, 2, 3)

states <- sample(possible.states, 252*252, replace=TRUE)

rpsmat <- matrix(states, nrow=252, ncol=252)

colnames(rpsmat) <- paste("c", 1:252, sep="")
rownames(rpsmat) <- paste("r", 1:252, sep="")
```

do one generation of updates - update matrix and frequencies of each cell type
global

```{r, eval=TRUE}
fc <- sum(rpsmat==1)/(nrow(rpsmat)*ncol(rpsmat))
fr <- sum(rpsmat==2)/(nrow(rpsmat)*ncol(rpsmat))
fs <- sum(rpsmat==3)/(nrow(rpsmat)*ncol(rpsmat))

```

local - ID point, then get the frequencies
assume that the updating for an empty cell depends on the absolute frequency of the other cells - not their relative frequencies
```{r}
column <- sample(2:251, 1)
row <- sample(2:251, 1)

focal <- rpsmat[row, column]

rpsmat[(row-1):(row+1), (column-1): (column+1)]

# frequency of C in local environment

c.local <- function(X, i, j) {sum(X[(i-1):(i+1),(j-1):(j+1)]==1)}/9
r.local <- function(X, i, j) {sum(X[(i-1):(i+1),(j-1):(j+1)]==2)}/9
s.local <- function(X, i, j) {sum(X[(i-1):(i+1),(j-1):(j+1)]==3)}/9

fc.loc <- c.local(rpsmat, row, column)

fr.loc <- r.local(rpsmat, row, column)

fs.loc <- s.local(rpsmat, row, column)

update <- function(matrix, row, column, prob.e, prob.s){
if(focal==0){
  matrix[row,column] = sample(c(1,2,3), 1, prob=prob.e, replace=TRUE)
  } else if(focal==1) {
    matrix[row, column] = sample(c(0,1), 1, prob=c(delta.c, 1-delta.c), replace=TRUE)
     } else if(focal==2){
       matrix[row, column] = sample(c(0,2), 1, prob=prob.s, replace=TRUE)
       } else if (focal==3) matrix[row, column] = sample(c(0,3), 1, prob=c(delta.r, 1-delta.r), replace=TRUE)
return(matrix[row, column])
}  

rpsmat[row, column] <- update(rpsmat, row, column, c(fc, fr, fs), c(delta.s, 1-delta.s))


# for loop

ngen=100000

abundance.c <- c()
abundance.s <- c()
abundance.r <- c()



for(k in 1:ngen){
  
  column <- sample(2:251, 1)
row <- sample(2:251, 1)

#frequency of each cell type in global environment
fc <- sum(rpsmat==1)/(nrow(rpsmat)*ncol(rpsmat))
fr <- sum(rpsmat==2)/(nrow(rpsmat)*ncol(rpsmat))
fs <- sum(rpsmat==3)/(nrow(rpsmat)*ncol(rpsmat))

  # death probability of susceptibles
  delta.s <- delta.s0 + tau*fc

# frequencies for local neighborhood
fc.loc <- sum(rpsmat[(row-1):(row+1), (column-1):(column+1)]==1)/9

fr.loc <- sum(rpsmat[(row-1):(row+1), (column-1):(column+1)]==2)/9

fs.loc <- sum(rpsmat[(row-1):(row+1), (column-1):(column+1)]==3)/9

focal <- rpsmat[row, column]

rpsmat[row, column] <- update(rpsmat, row, column, c(fc, fr, fs), c(delta.s, 1-delta.s))


abundance.c[k] <- sum(rpsmat==1)
abundance.r[k] <- sum(rpsmat==2)
abundance.s[k] <-  sum(rpsmat==3)

}

plot(x=1:100000, y=seq(from=9.5, to=10, by=0.5/99999), type="n")
  
lines(log(abundance.r), col="green")
lines(log(abundance.s), col="blue")
lines(log(abundance.c), col="red")

test <- rpsmat[1:10,1:10]
```


test <- rpsmat[1:10,1]


for loop for updating

save frequencies at each generation

 figure out how to save matricies every N generations (in a list?)

 work with neighborhood sizes 

```{r}
ngen=1000

abundance.c <- c()
abundance.s <- c()
abundance.r <- c()

test <- rpsmat[1:10,1:10]

for(k in 1:ngen){
  
  column <- sample(2:9, 1)
row <- sample(2:9, 1)

  #frequency of each cell type in global environment
  fc <- sum(test==1)/(nrow(test)*ncol(test))
fr <- sum(test==2)/(nrow(test)*ncol(test))
fs <- sum(test==3)/(nrow(test)*ncol(test))

  # death probability of susceptibles
  delta.s <- delta.s0 + tau*fc


fc.loc <- sum(test[(row-1):(row+1), (column-1):(column+1)]==1)/9

fr.loc <- sum(test[(row-1):(row+1), (column-1):(column+1)]==2)/9

fs.loc <- sum(test[(row-1):(row+1), (column-1):(column+1)]==3)/9



focal <- test[row, column]

test[row, column] <- update(test, row, column, c(fc, fr, fs), c(delta.s, 1-delta.s))

abundance.c[k] <- sum(test==1)
abundance.r[k] <- sum(test==2)
abundance.s[k] <-  sum(test==3)
}

plot(x=1:1000, y=seq(from=1, to=5, by=4/999), type="n")
  
lines(log(abundance.r), col="green")
lines(log(abundance.s), col="blue")
lines(log(abundance.c), col="red")


test


update <- function(matrix, row, column, size){
if(size == "local"){prob.e <- c(fc, fr, fs)} else prob.e <- c(fc.loc, fr.loc, fs.loc)
if(focal==0){
  matrix[row,column] = sample(c(1,2,3), 1, prob=prob.e, replace=TRUE)
  } else if(focal==1) {
    matrix[row, column] = sample(c(0,1), 1, prob=c(delta.c, 1-delta.c), replace=TRUE)
     } else if(focal==2){
       matrix[row, column] = sample(c(0,2), 1, prob=prob.s, replace=TRUE)
       } else if (focal==3) matrix[row, column] = sample(c(0,3), 1, prob=c(delta.r, 1-delta.r), replace=TRUE)
return(matrix[row, column])
}  

update(test, 7, 7, "local")



```
