---
title: "The Maintenance of Biodiversity in a Dispersal-Limited Strains of E. coli"
author: "Edna Chiang and Marian Schmidt"
date: "November 24, 2014"
output: html_document
---


##The Problem
1. How do local dispersal and species interactions influence biodiversity? 
2. How do global dispersal and species interactions influence biodiversity?


##Our Assumptions
* Simple community that perfectly follows Rock-Paper-Scissors model:
      + Resistant *E.coli* (R) beats colicin-producing *E.coli* (C)
      + Sensitive *E.coli* (S) beats R 
      + C beats S 
* No top-down or bottom up controls on community.
      + No outside disturbance.
      + No spatial or temporal environmental heterogeneity.
* Assume 2 extremes:
      + Local includes only adjacent cells.  Location is everything.
      + Global includes entire community.  All interactions created equally, location doesn't matter.
* Assumes background death rates are constant for each strain.
* Assumes no evolution of each strain (e.g. no evolution of S to R).


##Reproduction of Figure 1a-d

```{r, echo=FALSE, fig.height = 5, fig.width = 10}
source('Functions05.R')

########  RUN LOCAL SIMULATIONS 
conditions <- c("S", "R", "C", "E")
set.seed(1)
max <- createMatrix(conditions, nrows = 50, ncols = 50)
bigmax <- runLocalSims2(max)

########  RUN GLOBAL SIMULATIONS 
cultbigmax <- runGlobalSims2(max) # cult = culture, we're going global

#### MAKE PLOT
par(mfrow=c(1,2))
plotLogTime(bigmax, "Local")
plotLogTime(cultbigmax, "Global")
```





##Experiment
As you can see above, the resistant *E.coli* (R) wins the global scenario.  This is because R has a slighly lower death rate of 10/32 than the colicin-producing *E.coli* (C), which has a death rate of 1/3.  

**Research questions:**  
1.  If R and C have the same death rate, do R, C, & S coexist in the **local** simulations?  
2.  If R and C have the same death rate, do R, C & S coexist in the **global** simulations?  
**Hypotheses:**   
1.  R, C, and S will coexist if R & C have the same death rate in **local** simulations.  
2.  R and C will coexist and cause S to go extinct if R & C have the same death rate in **global** simulations.   


```{r, echo = FALSE}
########  RUN LOCAL SIMULATIONS 
#expmax <- runLocalSims2(max, Rdeath3 = 1/3)
#plotLogTime(expmax, "Local")

########  RUN GLOBAL SIMULATIONS 
#expbigmax <- runGlobalSims2(max, Rdeath = 1/3) # cult = culture, we're going global
###### MAKE LINE PLOT FOR GLOBAL
#plotLogTime(expbigmax, "Global")
```


#Discussion of Our Experiment





## TIME TO GIF IT OUT  

```{r, echo = FALSE, fig.height = 10, fig.width = 10, fig.show = 'animate'}
gif(bigmax)
```

```{r, echo = FALSE, fig.height = 5, fig.width = 5, fig.show = 'animate'}
gif(cultbigmax)
```

##GIF

