---
title: "assignment_5"
author: "Iris"
date: "November 7, 2014"
output: html_document
---
death probabilities

```{r, eval=TRUE}
delta.c <- 1/3
delta.r <- 10/32
tau <- 3/4
delta.s0 <- 1/4
delta.s <- delta.s0 + tau*fc
```

 put out initial outlay - start at equal abundances (from Fig.1 line graphs)
 
 0 = empty
 1=C
 2=R
 3=S
 
```{r}
possible.states <- c(0, 1, 2, 3)

states <- sample(possible.states, 252*252, replace=TRUE)

rpsmat <- matrix(states, nrow=252, ncol=252)

colnames(rpsmat) <- paste("c", 1:252, sep="")
rownames(rpsmat) <- paste("r", 1:252, sep="")
```

do one generation of updates - update matrix and frequencies of each cell type
global

```{r, eval=TRUE}
fc <- sum(rpsmat==1)/(nrow(rpsmat)*ncol(rpsmat))
fr <- sum(rpsmat==2)/(nrow(rpsmat)*ncol(rpsmat))
fs <- sum(rpsmat==3)/(nrow(rpsmat)*ncol(rpsmat))

```

local - ID point, then get the frequencies
assume that the updating for an empty cell depends on the absolute frequency of the other cells - not their relative frequencies
```{r}
column <- sample(2:251, 1)
row <- sample(2:251, 1)

focal <- rpsmat[row, column]

rpsmat[(row-1):(row+1), (column-1): (column+1)]

# frequency of C in local environment

c.local <- function(X, i, j) {sum(X[(i-1):(i+1),(j-1):(j+1)]==1)}/9
r.local <- function(X, i, j) {sum(X[(i-1):(i+1),(j-1):(j+1)]==2)}/9
s.local <- function(X, i, j) {sum(X[(i-1):(i+1),(j-1):(j+1)]==3)}/9

fc.loc <- c.local(rpsmat, row, column)

fr.loc <- r.local(rpsmat, row, column)

fs.loc <- s.local(rpsmat, row, column)

update <- function(matrix, row, column, prob.e, prob.s){
if(focal==0){
  matrix[row,column] = sample(c(1,2,3), 1, prob=prob.e, replace=TRUE)
  } else if(focal==1) {
    rpsmat[row, column] = sample(c(0,1), 1, prob=c(delta.c, 1-delta.c), replace=TRUE)
     } else if(focal==2){
       rpsmat[row, column] = sample(c(0,2), 1, prob=prob.s, replace=TRUE)
       } else if (focal==3) rpsmat[row, column] = sample(c(0,3), 1, prob=c(delta.r, 1-delta.r), replace=TRUE)
return(matrix[row, column])
}  

rpsmat[row, column] <- update(rpsmat, row, column, c(fc, fr, fs), c(delta.s, 1-delta.s))


```





for loop for updating

save frequencies at each generation

 figure out how to save matricies every N generations (in a list?)

 work with neighborhood sizes 


