Assignment 5
============

### The problem...
One problem that confounds biologists is how best to explain the
biodiversity that surrounds us.  For example, there are antibiotic producing
and sensitive strains of the same species that reside in the same
populations.  Also, there are phage that stably co-exist with their
bacterial prey.  These types of relationships were studied in a [paper published
in the journal Nature](http://www.nature.com/nature/journal/v418/n6894/full/nature00823.html) using the model of [Rock-Paper-Scissors](http://en.wikipedia.org/wiki/Rock-paper-scissors).

### The assignment...
You will quickly notice that the authors did not make their code publicly
available and do not detail all of their assumptions. Your job for this
assignment is to attempt to recreate the line graphs shown in Figure 1 from
either of these two papers.  
At a minimum, you need to create functions that
generate the starting matrix and a 
function that uses the starting matrix to
runs the simulation for a specified number of generations.  
You also need to
make the functions flexible enough to allow the user to alter one or more set of
variables. As output, you need to generate a knitr document that describes...

* the problem
* your assumptions
* a reproduction of their line graphs
* an experiment where you outline your hypothesis and what you learned from the experiment.
* if you can figure out how to create time-lapsed heatmaps of the state of the
population (also shown in Fig. 1) as a GIF, I will give you a considerable
amount of extra credit
* you may no more than one for/repeat/while loop (although none would be even better)

### Game plan...
Complete the exercise and submit it as a pull request to the [Assignment 5 repository](https://github.com/microbialinformatics/assignment05). You should not use any packages beyond the base R system. This assignment is due on Wednesday, November 26th. Be sure to include all R, Rmd, and md files in your commits. There are two approaches you and your partner can take to work on this project:

* Sit at the same computer and work together and have one person make all of the commits and pushes to GitHub.
* Divide and conquer the functions with one person creating the repository and make the other person a collaborator. Synthesize your code using pull requests
*try options 1 and 2 and if that fails, do project yourself.

**Problem
The authors wanted to model how competing species can exist together in a non-transitive community. They assessed the how local and global community dynamics influences population structure by competing three types of E. coli and measuring abundances of each type.  They tested the prediction that local interaction and dispersal are sufficient for all types in a community to exist, but when dispersal, movement, and interaction occur over a larger scale, diversity is lost.

**Assumptions
Assumptions used in this paper 
1. initially community structure is random (empty, or C,S,R types are equally likely to initially occupy a niche (cell) in the community)
2. Any point occupied by empty, C,S,or R can be changed
3. Changes in given locations are dependent on the current occupancy as well as local occupancy (neighbors or global community) with discinct probabilities relating to cell types.
4. These probabilities are:
S-->0  :  delta_S = delta_S,0 + tau * f_C
S-->S :  1 - delta_S
R-->0 :  delta_R
R-->R :  1 - delta_R
C-->0 :  delta_C
C-->C : 1 - delta_C
0-->S : f_S
0-->R : f_R
0-->C : f_C
0-->0 : 1 - f_S - f_R - f_C
5. Under these probabilities a rock-paper-scissors dynamic occurs where: S displaces R because S grows better, R displaces C because R grows better, and C can displace S because C kills S. 
6. neighborhoods include 8 cells surrounding a point, and neighborhoods of cells on the edge include distant cells (furthest away from index cell in the same plane)

**Experiment
The authors initially have empty cells in their matrix. Some communities are saturated. I wanted to know how the population dynamics would change if the starting community were fully saturated with S, C or R cells. I hypothesize that the S cells would die out faster in this scenario because there are no empty cells to be randomly changed to occupied by other cell types, so the overall displacement would play out faster. 
Results line graph below.
interpretation: the abundences are lower and the cells do not appear to change. Perhaps with no empty cells the cell types cannot change from their initial seeding density/locations because there is no room to increase or decrease the abundance of any of the three cell types. An alternative interpretation is that I have an error in my coding and the experimental competion matrix does not update for each round in the forloop.


```{r}

#initial set-up to start simulations
#fill matrix with randomly selected 4 choices
#C=1,S=2,R=3, E=4 empty (4 choices)
seeds<-sample(1:4,2500, replace=T) #pick the random numbers to fill matrix
initial.matrix<-matrix(seeds,nrow=50, ncol=50) #create matrix with random numbers
#pick random cell in matrix to change
#function takes initial random cell from matrix and changes it based on local neighborhood dynamics
local.comp<-function(initial.matrix) { 
r<-initial.matrix[(sample(1:50,1, replace=T))]
c<-initial.matrix[(sample(1:50,1, replace=T))]
#select local neighborhood based on index
#local neighborhood is 3 above, 2 each side, 3 on bottom of index
 #rows:
#a center upper
  up.row<-r+1
  if(up.row>50){up.row<-1}
  else if(up.row<1){up.row<-50}
  # b center lower
  down.row<-r-1
  if(down.row>50){down.row<-1}
  else if(down.row<1){down.row<-50}
  #c right upper corner
  r.up.corner<-r+1
  if(r.up.corner>50){r.up.corner<-1}
  else if(r.up.corner<1){r.up.corner<-50}
  #d left upper corner
  l.up.corner<-r+1
 if(l.up.corner>50){l.up.corner<-1}
  else if(l.up.corner<1){l.up.corner<-50}
  #e center right
  center.r<-r
 if(center.r>50){center.r<-1}
  else if(center.r<1){center.r<-50}
  # f center left
  center.l<-r
 if(center.l>50){center.l<-1}
  else if(center.l<1){center.l<-50}
  #g right lower corner
  r.low.corner<-r-1
if(r.low.corner>50){r.low.corner<-1}
  else if(r.low.corner<1){r.low.corner<-50}
  #h left lower corner 
  l.low.corner<-r-1
if(l.low.corner>50){l.low.corner<-1}
  else if(l.low.corner<1){l.low.corner<-50}
   #Columns:
  #a upper
  up.col<-c
if( up.col>50){up.col<-1}
  else if( up.col<1){up.col<-50}
  # b lower
  low.col<-c
if( low.col>50){low.col<-1}
  else if( low.col<1){low.col<-50}
  # c right upper corner
  r.up.cor.col<-c+1
if( r.up.cor.col>50){r.up.cor.col<-1}
  else if( r.up.cor.col<1){r.up.cor.col<-50}
  # d left upper corner
  l.up.cor.col<-c-1
if( l.up.cor.col>50){l.up.cor.col<-1}
  else if( l.up.cor.col<1){l.up.cor.col<-50}
  # e center right
  center.r.col<-c+1
if( center.r.col>50){center.r.col<-1}
  else if( center.r.col<1){center.r.col<-50}
  # f center left
  center.l.col<-c-1
if( center.l.col>50){center.l.col<-1}
  else if( center.l.col<1){center.l.col<-50}
  # g right lower corner
  r.lower.cor.col<-c+1
if( r.lower.cor.col>50){r.lower.cor.col<-1}
  else if( r.lower.cor.col<1){r.lower.cor.col<-50}
  # h left lower corner 
  l.lower.cor.col<-c-1
if( l.lower.cor.col>50){l.lower.cor.col<-1}
  else if( l.lower.cor.col<1){l.lower.cor.col<-50}
  # put together neighbor indices
  A<-initial.matrix[up.row,up.col]
  B<-initial.matrix[down.row,low.col]
  C<-initial.matrix[r.up.corner, r.up.cor.col]
  D<-initial.matrix[l.up.corner, l.up.cor.col]
  E<-initial.matrix[center.r, center.r.col]
  G<-initial.matrix[center.l, center.l.col]
  H<-initial.matrix[ r.low.corner,r.lower.cor.col]
  I<-initial.matrix[l.low.corner, l.lower.cor.col]
#if the cell is not empty
  #change based on given probabilities:#C=1/3 C is 1
  #S0=1/4+(3/4*fractionC) s is 2
  #R is 10/32 R is 3
  #tau =3/4 and is used in delta S calculation
  #start with vector of neighbors
    make.vec<-c(A,B,C,D,E,G,H,I)
  #find how many Cs there are in neighborhood
  #calculate fraction of C to use in delta_S calculation
    ones<-length(grep(1,make.vec))
    fract.ones<-ones/length(make.vec)
#change based on the given probabilities. Cs= cell sample scheme 
  #C=1
  if(initial.matrix[r,c]==1) {newcell<-sample((c(4,1)),1,prob=(c((1/3),(1-1/3))))}
  #S=2
  else if(initial.matrix[r,c]==2) {newcell<-sample((c(4,2)),1,prob=(c(((.25+.75*fract.ones)),(1-(.25+(.75*fract.ones))))))}
  #R=3
  else if(initial.matrix[r,c]==3) {newcell<-sample((c(3,4)),1,prob=(c((1-10/32),(10/32))))}
#change the type based on what is in neighborhood if cell is empty
  #probability of it becoming 1,2,3,0 is based on probability 
  else{
    newcell<-make.vec[sample(1:8,1)]} 
initial.matrix[r,c]<<-newcell
#reassign initial.matrix for next round
#return matrix
   }
##############################################
#Sampling for global competition
global.comp<-function(initial.matrix) {
  #creating random cell to pick for type to change to if cell picked is empty
  new.row<-sample(1:50,1)
  new.col<-sample(1:50,1)
  #pick the cell to change
r<-initial.matrix[(sample(1:50,1, replace=T))]
c<-initial.matrix[(sample(1:50,1, replace=T))]
#if the cell is not empty
  #change based on given probabilities:#C=1/3 C is 1
  #S0=1/4+(3/4*fractionC) s is 2
  #R is 10/32 R is 3
  #tau =3/4 and is used in delta S calculation
 #find how many Cs there are in neighborhood
  #calculate fraction of C to use in delta_S calculation
    ones<-length(grep(1,initial.matrix))
    fract.ones<-ones/length(initial.matrix)
#change based on the given probabilities. Cs= cell sample scheme 
  #C=1
  if(initial.matrix[r,c]=="1") {newcell<-sample((c(4,1)),1,prob=(c((1/3),(1-1/3))))}
  #S=2
  else if(initial.matrix[r,c]=="2") {newcell<-sample((c(4,2)),1,prob=(c(((.25+.75*fract.ones)),(1-(.25+(.75*fract.ones))))))}
  #R=3
  else if(initial.matrix[r,c]=="3") {newcell<-sample((c(3,4)),1,prob=(c((1-10/32),(10/32))))}
#change the type based on what is in neighborhood if cell is empty
  #probability of it becoming 1,2,3,0 is based on probability 
  else {newcell<-initial.matrix[new.row,new.col]}
  initial.matrix[r,c]<<-newcell
}
#EXPERIMENTAL: non-wrapping matrix for local competition
#make matrix for no empties experiment 
ex.seeds<-sample(1:3,2500, replace=T) #pick the random numbers to fill matrix
initial.matrix<-matrix(ex.seeds,nrow=50, ncol=50) #create matrix with random numbers
experimental.comp<-function(initial.matrix) {
r<-initial.matrix[(sample(1:50,1, replace=T))]
c<-initial.matrix[(sample(1:50,1, replace=T))]
#select local neighborhood based on index
#local neighborhood is 3 above, 2 each side, 3 on bottom of index
 #rows:
#a center upper
  up.row<-r+1
  if(up.row>50){up.row<-1}
  else if(up.row<1){up.row<-50}
  # b center lower
  down.row<-r-1
  if(down.row>50){down.row<-1}
  else if(down.row<1){down.row<-50}
  #c right upper corner
  r.up.corner<-r+1
  if(r.up.corner>50){r.up.corner<-1}
  else if(r.up.corner<1){r.up.corner<-50}
  #d left upper corner
  l.up.corner<-r+1
 if(l.up.corner>50){l.up.corner<-1}
  else if(l.up.corner<1){l.up.corner<-50}
  #e center right
  center.r<-r
 if(center.r>50){center.r<-1}
  else if(center.r<1){center.r<-50}
  # f center left
  center.l<-r
 if(center.l>50){center.l<-1}
  else if(center.l<1){center.l<-50}
  #g right lower corner
  r.low.corner<-r-1
if(r.low.corner>50){r.low.corner<-1}
  else if(r.low.corner<1){r.low.corner<-50}
  #h left lower corner 
  l.low.corner<-r-1
if(l.low.corner>50){l.low.corner<-1}
  else if(l.low.corner<1){l.low.corner<-50}
   #Columns:
  #a upper
  up.col<-c
if( up.col>50){up.col<-1}
  else if( up.col<1){up.col<-50}
  # b lower
  low.col<-c
if( low.col>50){low.col<-1}
  else if( low.col<1){low.col<-50}
  # c right upper corner
  r.up.cor.col<-c+1
if( r.up.cor.col>50){r.up.cor.col<-1}
  else if(r.up.cor.col<1){r.up.cor.col<-50}
  # d left upper corner
  l.up.cor.col<-c-1
if( l.up.cor.col>50){l.up.cor.col<-1}
  else if( l.up.cor.col<1){l.up.cor.col<-50}
  # e center right
  center.r.col<-c+1
if( center.r.col>50){center.r.col<-1}
  else if( center.r.col<1){center.r.col<-50}
  # f center left
  center.l.col<-c-1
if( center.l.col>50){center.l.col<-1}
  else if( center.l.col<1){center.l.col<-50}
  # g right lower corner
  r.lower.cor.col<-c+1
if( r.lower.cor.col>50){r.lower.cor.col<-1}
  else if( r.lower.cor.col<1){r.lower.cor.col<-50}
  # h left lower corner 
  l.lower.cor.col<-c-1
if( l.lower.cor.col>50){l.lower.cor.col<-1}
  else if( l.lower.cor.col<1){l.lower.cor.col<-50}
  # put together neighbor indices
  A<-initial.matrix[up.row,up.col]
  B<-initial.matrix[down.row,low.col]
  C<-initial.matrix[r.up.corner, r.up.cor.col]
  D<-initial.matrix[l.up.corner, l.up.cor.col]
  E<-initial.matrix[center.r, center.r.col]
  G<-initial.matrix[center.l, center.l.col]
  H<-initial.matrix[ r.low.corner,r.lower.cor.col]
  I<-initial.matrix[l.low.corner, l.lower.cor.col]
#if the cell is not empty
  #change based on given probabilities:#C=1/3 C is 1
  #S0=1/4+(3/4*fractionC) s is 2
  #R is 10/32 R is 3
  #tau =3/4 and is used in delta S calculation
  #start with vector of neighbors
    make.vec<-c(A,B,C,D,E,G,H,I)
  #find how many Cs there are in neighborhood
  #calculate fraction of C to use in delta_S calculation
    ones<-length(grep(1,make.vec))
    fract.ones<-ones/length(make.vec)
#change based on the given probabilities. Cs= cell sample scheme 
  #C=1
  if(initial.matrix[r,c]==1) {newcell<-sample((c(4,1)),1,prob=(c((1/3),(1-1/3))))}
  #S=2
  else if(initial.matrix[r,c]==2) {newcell<-sample((c(4,2)),1,prob=(c(((.25+.75*fract.ones)),(1-(.25+(.75*fract.ones))))))}
  #R=3
  else if(initial.matrix[r,c]==3){newcell<-sample((c(3,4)),1,prob=(c((1-10/32),(10/32))))}
initial.matrix[r,c]<<-newcell}
#reassign initial.matrix for next round
#return matrix

```
#results


```{r}
#forloop for local.comp
fill<-c("NA")
empty.matrix<-matrix(fill,nrow=2500,ncol=1001) #make empty matrix to fill 
update.matrix<-initial.matrix #call input matrix something else to update it
j<-1
for( i in 1:((2500)*1000)) { #for each update
  if (i==1 | i %% 2500==0){ #
    empty.matrix[,j]<-update.matrix
    j <-j+1
  }
  update.matrix<-local.comp(initial.matrix)
  }
colnames(empty.matrix)<-1:1001
return(empty.matrix)

#get abundances for local comp
#C,S,R
frequencies<-function(empty.matrix){
  empty.matrix.df<-as.data.frame(empty.matrix)
  label.matrix<-sapply(empty.matrix.df, function(x) table(factor(x, levels=(c("1","2","3","4")))))
  lable.matrix.log<-log10(label.matrix)
  lable.matrix.log[is.infinite(lable.matrix.log)] = 0
  lable.matrix.log<-as.data.frame(lable.matrix.log)
  return(lable.matrix.log)
  }

freq.matrix<-frequencies(empty.matrix)                    
#transpose
freq.matrix.t<-as.data.frame(t(freq.matrix))
plot(freq.matrix.t$"1",type="l", col="green",ylim=(c(0,5)),ylab="Log(abundance)",xlab="Time",main="Local neighbourhood", xlim=c(0,ncol(freq.matrix)))
lines(freq.matrix.t$"2", type="l", col="red")
lines(freq.matrix.t$"3", type="l", col="blue")

############################################
#global competition
fill<-(c("NA"))
empty.matrix<-matrix(fill,nrow=2500,ncol=501) #make empty matrix to fill 
update.matrix<-initial.matrix #call input matrix something else to update it
j<-1
for( i in 1:((2500)*500)) { #for each update
  if (i==1 | i %% 2500==0){ #
    empty.matrix[,j]<-update.matrix
    j <-j+1
  }
  update.matrix<-global.comp(initial.matrix)
  }
colnames(empty.matrix)<-1:501
return(empty.matrix)


frequencies<-function(empty.matrix){
  empty.matrix.df<-as.data.frame(empty.matrix)
  label.matrix<-sapply(empty.matrix.df, function(x) table(factor(x, levels=(c("1","2","3","4")))))
  lable.matrix.log<-log10(label.matrix)
  lable.matrix.log[is.infinite(lable.matrix.log)] = 0
  lable.matrix.log<-as.data.frame(lable.matrix.log)
  return(lable.matrix.log)
  }

freq.matrix<-frequencies(empty.matrix)                    
#transpose
freq.matrix.t<-as.data.frame(t(freq.matrix))
plot(freq.matrix.t$"1",type="l", col="green",ylim=(c(0,5)),ylab="Log(abundance)",xlab="Time",main="Global neighbourhood", xlim=c(0,ncol(freq.matrix.t)))
lines(freq.matrix.t$"2", type="l", col="red")
lines(freq.matrix.t$"3", type="l", col="blue")

#############
#loop for experimental
#make matrix without empty cells
ex.seeds<-sample(1:3,2500, replace=T) #pick the random numbers to fill matrix
initial.matrix<-matrix(ex.seeds,nrow=50, ncol=50) #create matrix with random numbers
fill<-(c("NA"))
empty.matrix<-matrix(fill,nrow=2500,ncol=1001) #make empty matrix to fill 
update.matrix<-initial.matrix #call input matrix something else to update it
j<-1
for( i in 1:((2500)*1000)) { #for each update
  if (i==1 | i %% 2500==0){ #
    empty.matrix[,j]<-update.matrix
    j <-j+1
  }
  update.matrix<-local.comp(initial.matrix)
  }
colnames(empty.matrix)<-1:1001
return(empty.matrix)

#plot experiment
freq.matrix<-frequencies(empty.matrix)                    
#transpose
freq.matrix.t<-as.data.frame(t(freq.matrix))
plot(freq.matrix.t$"1",type="l", col="green",ylim=(c(0,5)),ylab="Log(abundance)",xlab="Time",main="Experiment: Local neighbourhood, no empty slots", xlim=c(0,ncol(freq.matrix)))
lines(freq.matrix.t$"2", type="l", col="red")
lines(freq.matrix.t$"3", type="l", col="blue")


```


#sources
http://stackoverflow.com/questions/9109778/random-sampling-matrix
help from Danielle and her group
searching github/micro 612 class website
google
