---
title: "Report--Assignment05"
author: "Caroline Van Steendam, Hui Zhang"
date: "Friday, November 21, 2014"
output: html_document
---

***

### Problem

Here is a system involves a community of three competing species: colicinogenic E.coli (C), colicin-sensitive bacteria (S) and colicin-resistant bacteria (R). And this C-S-R community satisfies a relationship similar to rock-paper-scissors game: S can displace R, R can displace C and C can displace S. The problem is how would this relationship affects the biodiveristiy in a small spatial scale and in a large spatial scale ecological processes.



### Assumptions
 * Assume a 50*50 lattice as a community
 * Each lattice point is randomly assigned by R, C, S cell and E (empty state)
 * Small spatial scale ecological processes (local neighborhood) is the eight lattice points around the specific focal point
 * Large spatical scale ecological processes (global neighborhood) is defined as all the lattice points in the community except the specific focal point
 * Update rules: 
 The update state decpends on the current state and its neighborbood state. Rules are the same for both small scale and large scale ecological processes.
  

State before update | State after update | Probability
----------- | ------------ | ----------- 
E | S | f_S
E | R | f_R
E | C | f_C
E | E | f_E
S | E | delta_S
S | S | 1-delta_S
R | E | delta_R
R | R | 1-delta_R
C | E | delta_C
C | C | 1-delta_C

* NOTE
     + f_S, f_R, f_C and f_E are the fractions of S, R, C and E in the neighborhood of update point.
     + delta_S, delta_R, and delta_C are the probability of S, R, and C cells to be killed.
     + delta_S = delta_S0 + tau*f_C     
       delta_S0 is the probability of S death without C, delta_S0 = 1/4     
       tau is the toxicity of C to S cells, tau = 3/4
     + delta_R = 10/32
     + delta_C = 1/3


### Reproduction of line graphs

```{r, echo=FALSE}
rm(list=ls())
#Creation of grid
create_grid <- function(size){
  types <- c("E", "R", "C", "S")
  data <- sample(types, size*size, replace = T)
  grid <- matrix(data, size, size)

return(grid)
}

#Setting up local neighbourhood and calculating fractions
find_local <- function (focalpoint_row, focalpoint_column, grid, size){  
    
#When the focal point is found somehwere in the middle,
if(focalpoint_row != 1 && focalpoint_row != size){
    local <- grid[(focalpoint_row-1):(focalpoint_row+1),]
    if (focalpoint_column != 1 && focalpoint_column != size){
      local <- local[, (focalpoint_column-1):(focalpoint_column+1)]
      local[2, 2] <- NA
      divider <- 8}      
    else if (focalpoint_column == size){
      local <- local[, (focalpoint_column-1):(focalpoint_column)]
      local[2, 2] <- NA
      divider <- 5}
    else{
      local <- local[, (focalpoint_column):(focalpoint_column+1)]
      local[2, 1] <- NA
      divider <- 5}
}

# When the focal point is found on the first row, 
else if (focalpoint_row == 1){
    local <- grid[(focalpoint_row):(focalpoint_row+1),]
    if (focalpoint_column != 1 && focalpoint_column != size){
      local <- local[, (focalpoint_column-1):(focalpoint_column+1)]
      local[1, 2] <- NA
      divider <- 5}   
    else if (focalpoint_column == size){
      local <- local[, (focalpoint_column-1):(focalpoint_column)]
      local[1, 2] <- NA
      divider <- 3}
    else{
      local <- local[, (focalpoint_column):(focalpoint_column+1)]
      local[1, 1] <- NA
      divider <- 3}}

#When the focal point is found on the last row, 
else{
    local <- grid[(focalpoint_row-1):(focalpoint_row),]
    if (focalpoint_column != 1 && focalpoint_column != size){
      local <- local[, (focalpoint_column-1):(focalpoint_column+1)]
      local[2, 2] <- NA
      divider <- 5}
    else if (focalpoint_column == size){
      local <- local[, (focalpoint_column-1):(focalpoint_column)]
      local[2, 2] <- NA
      divider <- 3}
    else{
      local <- local[, (focalpoint_column):(focalpoint_column+1)]
      local[2, 1] <- NA
      divider <- 3}}

#Calculate fraction and abundance 
  f_r <- length(grep("R", local))/divider
  f_c <- length(grep("C", local))/divider
  f_s <- length(grep("S", local))/divider
  f_e <- 1-f_r-f_c-f_s
  R_ab <- length(grep("R", grid))
  C_ab <- length(grep("C", grid))
  S_ab <- length(grep("S", grid))
  
  info <- list(f_e, f_r, f_c, f_s, R_ab, C_ab, S_ab)
  return(info)
}


find_global <- function (focalpoint_row, focalpoint_column, grid, size){
  grid[focalpoint_row, focalpoint_column] <- NA
  f_r <- length(grep("R", grid))/(size*size-1)
  f_c <- length(grep("C", grid))/(size*size-1)
  f_s <- length(grep("S", grid))/(size*size-1)
  f_e <- 1-f_r-f_c-f_s
  R_ab <- length(grep("R", grid))
  C_ab <- length(grep("C", grid))
  S_ab <- length(grep("S", grid))
  
  info <- list(f_e, f_r, f_c, f_s, R_ab, C_ab, S_ab)
  return(info)
}


#Setting up parameters.
delta_r <- 10/32
delta_c <- 1/3
delta_s0 <-1/4
tau <- 3/4


replace_focal <- function(focalpoint, neighbourhood_info){
  if (focalpoint == "E"){
    #Update empty space based on the occurence of the different types in the neighbourhood.
    focalpoint <- sample(c("E", "R", "C", "S"), 1, replace = T, prob = c(neighbourhood_info[[1]], neighbourhood_info[[2]], neighbourhood_info[[3]], neighbourhood_info[[4]]))}
  #Kill microorganism, with a probability equal to the given delta's. 
  else if (focalpoint == "R"){
    focalpoint <- sample(c("E", "R"), 1, replace = T, prob = c(delta_r, (1-delta_r)))}
  else if (focalpoint == "C"){
    focalpoint <- sample(c("E", "C"), 1, replace = T, prob = c(delta_c, (1-delta_c)))}
  else {
    delta_s <- delta_s0 + tau*neighbourhood_info[[3]]
    focalpoint <- sample(c("E", "S"), 1, replace = T, prob = c(delta_s, (1-delta_s)))}
    
    return(focalpoint)
}



#Updating the grid. This happens randomly and 2500 updates are called an 'epoch'.
#Inputs: the type of interaction (either "local" or "global"), the grid you want to update,the size of this grid, and the amount of updates you want to occur
update_grid <- function(interaction, grid, size, updates){

  #Initializing
  epoch <- size*size
  counter <- 0
  epochcounter <- 0
  R_abundance <- 0
  C_abundance <- 0
  S_abundance <- 0
  
    
  #Updating the grid using two approachinges: local interactions, or global interactions. 
  for (i in 1:(epoch*updates)){
    counter <- counter + 1
  #Randomly select a location and obtain its value in both grids. 
  focalpoint_row = sample(1:nrow(grid), 1)
  focalpoint_column = sample(1:ncol(grid), 1)
  focalpoint <- grid[focalpoint_row, focalpoint_column]
  
  if (interaction == "local"){
    #Call upon find_local to find the probabilities.
    info <- find_local(focalpoint_row, focalpoint_column, grid, size)
    #Replace the focal points according to their neighbourhoods (call upon replace_focal).
    grid[focalpoint_row, focalpoint_column] <- replace_focal(focalpoint, info)}
  else{    
    #Call upon find_global to find the probabilities for both grids. 
    info <- find_global(focalpoint_row, focalpoint_column, grid, size)
    #Replace the focal points in both grids according to their neighbourhoods (call upon replace_focal).
    grid[focalpoint_row, focalpoint_column] <- replace_focal(focalpoint, info)}     

  #For each epoch, save the abundance information to generate the plots later on. This abundance information is already present in both info_local and info_global, so we only have to save this information in new vectors.
  if (counter == epoch) {
    epochcounter <- epochcounter + 1
    counter <- 0
        
    R_abundance[epochcounter] <- info[[5]]
    C_abundance[epochcounter] <- info[[6]]
    S_abundance[epochcounter] <- info[[7]]}

  output_update <- list(grid, R_abundance, C_abundance, S_abundance)
  return(output_update)
}
}


#plot the abundances of the local neighbourhoods. 
plot_abundance <- function(output_update, updates){
  x <- c(1:updates)
  #plot the abundance of R
  plot <- plot(x, log10(output_update[[2]]), ylim = c(0,10), type = "l", col = "green",xlab="Time",ylab="Log(abundance)")
  #plot the abundance of C
  points(x, log10(output_update[[3]]), type = "l", col = "red")
  #plot the abundance of S
  points(x, log10(output_update[[4]]), type = "l", col = "blue")
  #empty is not plotted
#   return(localplot)
}




## This doesn't work
#grid <- create_grid(50)
#output_update_local <- update_grid("local",grid,50,1000)
#plot_abundance(output_update_local,1000)
#output_update_global <- update_grid("global",grid,50,1000)
#plot_abundance(output_update_global,1000)


##Another way to output line graphs: save as image, and insert the image

```



### Experiment
(an experiment where you outline your hypothesis and what you learned from the experiment.)

  * Hypothesis    
Same community as before: a randomly assigned 50*50 lattice. All the parameters and assumptions are same as before, except the probability of R cells to be killed, delta_R, is changed to 3/4, which means R cell is much more easily to displaced by empty-state. Then in a well-mixed system (global neighborhood), R cell has less effect on C cell. C cell might be able to survive. Since there might be more empty-state lattice point, the probability of S cell to fill this point increase. So S cell might also be able to survive. We assume R, S and C cells could co-exist under this new conditions.
 
 
  * Global neighborhood plot
 
```{r,echo=FALSE}
#Creation of grid
create_grid <- function(size){
  types <- c("E", "R", "C", "S")
  data <- sample(types, size*size, replace = T)
  grid <- matrix(data, size, size)

return(grid)
}


find_global <- function (focalpoint_row, focalpoint_column, grid, size){
  grid[focalpoint_row, focalpoint_column] <- NA
  f_r <- length(grep("R", grid))/(size*size-1)
  f_c <- length(grep("C", grid))/(size*size-1)
  f_s <- length(grep("S", grid))/(size*size-1)
  f_e <- 1-f_r-f_c-f_s
  R_ab <- length(grep("R", grid))
  C_ab <- length(grep("C", grid))
  S_ab <- length(grep("S", grid))
  
  info <- list(f_e, f_r, f_c, f_s, R_ab, C_ab, S_ab)
  return(info)
}


update_grid <- function(grid, size, updates){

  #Initializing
  epoch <- size*size
  counter <- 0
  epochcounter <- 0
  R_abundance <- 0
  C_abundance <- 0
  S_abundance <- 0
  
    
  #Updating the grid using two approachinges: local interactions, or global interactions. 
  for (i in 1:(epoch*updates)){
    counter <- counter + 1
  #Randomly select a location and obtain its value in both grids. 
  focalpoint_row = sample(1:nrow(grid), 1)
  focalpoint_column = sample(1:ncol(grid), 1)
  focalpoint <- grid[focalpoint_row, focalpoint_column]
  
  grid[focalpoint_row, focalpoint_column] <- replace_focal(focalpoint, info)}     

  #For each epoch, save the abundance information to generate the plots later on. This abundance information is already present in both info_local and info_global, so we only have to save this information in new vectors.
  if (counter == epoch) {
    epochcounter <- epochcounter + 1
    counter <- 0
        
    R_abundance[epochcounter] <- info[[5]]
    C_abundance[epochcounter] <- info[[6]]
    S_abundance[epochcounter] <- info[[7]]}

  output_update <- list(grid, R_abundance, C_abundance, S_abundance)
  return(output_update)
 }



#plot the abundances of the local neighbourhoods. 
plot_abundance <- function(output_update, updates){
  x <- c(1:updates)
  plot <- plot(x, log10(output_update[[2]]), ylim = c(0,10), type = "l", col = "green",xlab="Time",ylab="Log(abundance)", main="Global Neighborhood")
  points(x, log10(output_update[[3]]), type = "l", col = "red")
  points(x, log10(output_update[[4]]), type = "l", col = "blue")
  #   return(localplot)
}
```
 

 * Discussion       
 




### Heatmap 
(if you can figure out how to create time-lapsed heatmaps of the state of the
population (also shown in Fig. 1) as a GIF, I will give you a considerable
amount of extra credit)
